---
title: "Homework_04"
author: "Madelynn Edwards"
date: "2024-02-07"
output: html_document
---

# Atomic Vectors and Basic R Lab

## Question #1
Suppose x = 1.1, a = 2.2, and b = 3.3. Assign each expression to the value of the variable z and print the value stored in z.

a. $x^{a^b}$ 
```{r}
x <- 1.1

a <- 2.2

b <- 3.3

z <- x^{a^b}
print(z)
```


b. (x^a^)^b^
```{r}
x <- 1.1

a <- 2.2

b <- 3.3

z <- (x^a)^b
print(z)

```

c. 3x^3^+2x^2^+1
```{r}
x <- 1.1

a <- 2.2

b <- 3.3

z <- (3*(x^3))+2*(x^2)+1
print(z)
```

## Question #2
Using the ```rep``` and ```seq``` functions, create the following vectors:

a. (1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)
```{r}
a <- seq(from = 1, to = 7, by = 1)

b <- seq(from = 8, to = 1, by = -1)

z <- c(a,b)
print(z)
```



b. (1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)
```{r}
my_vec<- c(1,2,3,4,5)
rep(x=my_vec, time=my_vec)
```

c. (5,4,4,3,3,3,2,2,2,2,1,1,1,1,1)
```{r}
my_vec<-c(5,4,3,2,1)
rep(x=my_vec, time= c(1,2,3,4,5))
```

## Question #3
Create a vector of two random uniform numbers. In a spatial map, these can be interpreted as x and y coordinates that give the location of an individual (such as a marked forest tree in a plot that has been mapped). Using one of R’s inverse trigonometry functions (```asin()```, ```acos()```, or ```atan()```), convert these numbers into polar coordinates

```{r}
set.seed(1000)
xy_coordinates <- runif(2)
polar_coordinates_r <- sqrt(xy_coordinates[1]+xy_coordinates[2])
print(polar_coordinates_r)
polar_coordinates_0 <- atan(xy_coordinates[2]/xy_coordinates[1])
print(polar_coordinates_0)
polar_coordinates <- c(polar_coordinates_r, polar_coordinates_0)
print(polar_coordinates)
```

## Question #4
Create a vector ```queue <- c("sheep", "fox", "owl", "ant")``` where queue represents the animals that are lined up to enter Noah’s Ark, with the sheep at the front of the line. Using R expressions, update ```queue``` as:

```{r}
queue <- c("sheep", "fox", "owl", "ant")

# a. the serpent arrives and gets in line;
queue <- c(queue, "serpent")
print (queue)

# b. the sheep enters the ark;
queue <- queue[-1]
print(queue)

# c. the donkey arrives and talks his way to the front of the line;
queue <- c("donkey", queue)
print(queue)

# d. the serpent gets impatient and leaves;
queue <- queue[-5]
print(queue)

# e. the owl gets bored and leaves;
queue <- queue[-3]
print(queue)

# f. the aphid arrives and the ant invites him to cut in line.
queue <- append(queue, "aphid", after=2)
print(queue)

# g. Finally, determine the position of the aphid in the line.
which(queue=="aphid")
```
## Question #5
Use R to create a vector of all of the integers from 1 to 100 that are not divisible by 2, 3, or 7.
```{r}

my_vec<- seq(1,100) # making vector sequence 1 to 100

print(my_vec)

#my_vec <- my_vec[which((my_vec%%2) >0)] # %% gives remainders after dividing, >=1 selecting those with a remainder greater than or equal to 1
#print(my_vec)
 
my_vec2<- my_vec[which((my_vec%%2)>=1 & (my_vec%%3)>=1 & (my_vec%%7)>=1)]

print(my_vec2)
```