---
title: "Homework_10"
author: "Madelynn Edwards"
date: "2024-04-03"
output: html_document
---

# For Loops and Randomization Tests

#### 1. Using a ```for loop```, write a function to calculate the number of zeroes in a numeric vector. Before entering the loop, set up a ```counter``` variable ```counter <- 0```. Inside the loop, add 1 to counter each time you have a zero in the vector. Finally, use ```return(counter)``` for the output.

```{r}
z <- rbinom(30, 1, 0.5)
z
counter <- 0
  
for (i in seq_along(z)){
    if(z[i] == 0) {
      counter <- counter + 1 
    }
  }
print(counter)

```


#### 2. Use subsetting instead of a loop to rewrite the function as a single line of code.

```{r}
length(z[z==0])



```


#### 3. Write a function that takes as input two integers representing the number of rows and columns in a matrix. The output is a matrix of these dimensions in which each element is the product of the row number x the column number.

```{r}

matrixfunction <- function(row = 2, col = 3){
    df <- expand.grid(x=1:row, y=1:col)
    print(df)
    mvec <- df$x*df$y
    #print(mvec)
    matrix <- matrix(mvec, nrow=row, ncol=col)
    print(matrix)
}

matrixfunction()
```


#### 4. Now let’s practice calling custom functions within a for loops. Use the code from previous lectures on loops and functions to complete the following steps:

I had some trouble with this question!

#### a. Simulate a dataset with 3 groups of data, each group drawn from a distribution with a different mean. The final data frame should have 1 column for group and 1 column for the response variable.
```{r}
n <- 15 # number of observations in each variable
var <- c("var1", "var2", "var3") # including three variable groups into the var object

data <- data.frame(matrix(nrow= n, ncol=2)) # make dataframe that will have two columns and 15 rows per variable
var_df <- NULL # empty object to assign temporary dataframe to

for(i in 1:length(var)) {
  data[1] <- rep(var[i], each=n) # repeating var (starting with var1) for first column
  data[2] <- rnorm(n=n, mean=runif(1)) # generating 15 random numbers for second column of dataframe, each with different mean for each var
  var_df <- rbind(var_df,data) # binding data in one dataframe
}

str(var_df)


```




#### b. Write a custom function that 1) reshuffles the response variable, and 2) calculates the mean of each group in the reshuffled data. Store the means in a vector of length 3.

```{r}
library(dplyr)

reshufflefunc <- function(n=15, v=3) {
  var_df[2] <- sample(var_df$X2, replace = FALSE) # reshuffling column data
  group_mean<- aggregate(x= var_df$X2, by = list(var_df$X1), FUN = mean)
  
  mean_vec <- group_mean$x # converting to vector
print(mean_vec)
    
}

reshufflefunc()
```


#### c.Use a for loop to repeat the function in b 100 times. Store the results in a data frame that has 1 column indicating the replicate number and 1 column for each new group mean, for a total of 4 columns.

```{r}

rep <- 1:100
df_rep <- NULL
mean_vec <- NULL

for (i in 1:length(rep)) {
  mean_vec <- reshufflefunc()
  var_df[3] <- mean_vec
  var_df[4] <- i
  df_rep <- rbind(df_rep, var_df)
}

str(df_rep)

```


#### d. Use qplot() to create a histogram of the means for each reshuffled group. Or, if you want a challenge, use ggplot() to overlay all 3 histograms in the same figure. How do the distributions of reshuffled means compare to the original means?

This doesnt seem right...

```{r}
library(ggplot2)


plot <- ggplot(data = df_rep)+
  aes(x=V3, fill=X1)+
  geom_histogram()
plot



```







