---
title: "Homework_08"
author: "Madelynn Edwards"
date: "2024-03-20"
output: html_document
---

# Data manipulations using the ```dplyr``` package

```{r warning=FALSE, message=FALSE}
library(tidyverse)
```

#### 1. Examine the structure of the ```iris``` data set. How many observations and variables are in the data set?

```{r warning=FALSE, message=FALSE}
data("iris")
class(iris) # data frame
tbl_df(iris) # convert to tibble
str(iris)
```

There are a total of 150 observations and 5 variables in the ```iris``` dataset.

#### 2. Create a new data frame ```iris1``` that contains only the species virginica and versicolor with sepal lengths longer than 6 cm and sepal widths longer than 2.5 cm. How many observations and variables are in the data set?

```{r}
 iris1 <- filter(iris, Species == "virginica" | Species == "versicolor", Sepal.Length > 6, Sepal.Width > 2.5)
 str(iris1)
```
There are now 56 observations in the new ```iris1``` dataframe and there are still 5 variables


#### 3. Now, create a ```iris2``` data frame from ```iris1``` that contains only the columns for Species, Sepal.Length, and Sepal.Width. How many observations and variables are in the data set?

```{r}
iris2 <- select(iris1, c(Species, Sepal.Length, Sepal.Width))
str(iris2)
```

In the ```iris2``` dataset, there are 56 observations and 3 variables.


#### 4. Create an ```iris3``` data frame from ```iris2``` that orders the observations from largest to smallest sepal length. Show the first 6 rows of this data set.

```{r}
iris3<-arrange(iris2, desc(Sepal.Length))
head(iris3)
```


#### 5. Create an ```iris4``` data frame from ```iris3``` that creates a column with a sepal area (length * width) value for each observation. How many observations and variables are in the data set?
```{r}
iris4<-mutate(iris3, Sepal.Area = Sepal.Length * Sepal.Width )
head(iris4)
str(iris4)
```
In the ```iris4``` dataset there are still 56 observations but now there are 4 variables (Sepal.Area added as the last column)



#### 6. Create ```iris5``` that calculates the average sepal length, the average sepal width, and the sample size of the entire ```iris4``` data frame and print ```iris5```.

```{r}
iris5<-summarize(iris4, MeanSepal.Length = mean(Sepal.Length), MeanSepal.Width = mean(Sepal.Width), number = n())
print(iris5)
```


#### 7. Finally, create ```iris6``` that calculates the average sepal length, the average sepal width, and the sample size for each species of in the ```iris4``` data frame and print ```iris6```.

```{r}
iris6 <- iris4%>%
  group_by(Species) %>%
  summarize(MeanSepal.Length = mean(Sepal.Length), MeanSepal.Width = mean(Sepal.Width), number = n())
print(iris6)

```


#### 8. In these exercises, you have successively modified different versions of the data frame ```iris1 iris2 iris3 iris4 iris5 iris6```. At each stage, the output data frame from one operation serves as the input fro the next. A more efficient way to do this is to use the pipe operator %>% from the tidyr package. See if you can rework all of your previous statements (except for ```iris5```) into an extended piping operation that uses iris as the input and generates ```irisFinal``` as the output.

```{r}
irisFinal <- iris %>%
  filter(Species == "virginica" | Species == "versicolor", Sepal.Length > 6, Sepal.Width > 2.5) %>%
  select(c(Species, Sepal.Length, Sepal.Width)) %>%
  arrange(desc(Sepal.Length)) %>%
  mutate(Sepal.Area = Sepal.Length * Sepal.Width) %>%
  group_by(Species) %>%
  summarize(MeanSepal.Length = mean(Sepal.Length), MeanSepal.Width = mean(Sepal.Width), number = n())
irisFinal

  


```

#### 9. Create a ‘longer’ data frame using the original iris data set with three columns named “Species”, “Measure”, “Value”. The column “Species” will retain the species names of the data set. The column “Measure” will include whether the value corresponds to Sepal.Length, Sepal.Width, Petal.Length, or Petal.Width and the column “Value” will include the numerical values of those measurements.

```{r}
iris_Longer <- iris %>% pivot_longer(cols = Sepal.Length:Petal.Width,
                      names_to = "Measure",
                      values_to = "Value")
head(iris_Longer)
                        
```









