---
title: "Homework_05"
author: "Madelynn Edwards"
date: "2024-02-14"
output: html_document
---

# Working with matrices, lists, and data frames

## Question #1

Assign to the variable ```n_dims``` a single random integer between 3 and 10.

```{r}
n_dims <- round(runif(min = 3, max = 10, 1))
print(n_dims)

```

* Create a vector of consecutive integers from 1 to n_dims2.
```{r}
consecutive_vector <- seq(1,n_dims^2)
consecutive_vector
```
* Use the sample function to randomly reshuffle these values.
```{r}
random <- sample(consecutive_vector)
random
```


* Create a square matrix with these elements.
```{r}
my_matrix <- matrix(data = random, ncol = 7, nrow=7)
```

* print out the matrix.
```{r}
my_matrix
```

* find a function in r to transpose the matrix.
```{r}
my_matrixT <- t(my_matrix)
```

* print it out again and note how it has changed.
```{r}
my_matrixT
```

* calculate the sum and the mean of the elements in the first row and then the last row.
```{r}
sum_r1 <- sum(my_matrixT[1,])
sum_r1
mean_r1 <- mean(my_matrixT[1,])
mean_r1

sum_r7 <- sum(my_matrixT[7,])
sum_r7
mean_r7 <- mean(my_matrixT[7,])
mean_r7
```

* read about the ```eigen()``` function and use it on your matrix
```{r}
em <- eigen(my_matrixT)

```


* look carefully at the elements of ```$values``` and ```$vectors``` in the output. What kind of numbers are these?

Imaginary

* dig in with the ```typeof()``` function to figure out their type.
```{r}
typeof(em)
typeof(em$values)
typeof(em$vectors)

```

* if have set your code up properly, you should be able to re-run it and create a matrix of different size because ```n_dims``` will change.

## Question #2

Create a list with the following named elements:

* ```my_matrix```, which is a 4 x 4 matrix filled with random uniform values
```{r}
my_matrix <- matrix(runif(16), ncol = 4, nrow = 4)
my_matrix
```


* ```my_logical``` which is a 100-element vector of TRUE or FALSE values. Do this efficiently by setting up a vector of random values and then applying an inequality to it.
```{r}
vector <- seq(1:100)
vector

my_logical <- (vector %% 7)>2
my_logical
```


* ```my_letters```, which is a 26-element vector of all the lower-case letters in random order.
```{r}
my_letters <- sample(letters[1:26])
my_letters
```

```{r}
my_list <- list(my_matrix, my_logical, my_letters)
my_list
```

Then, complete the following steps:

* create a new list, which has the element[2,2] from the matrix, the second element of the logical vector, and the second element of the letters vector.
```{r}
new_list <- list(my_list[[1]][2,2],my_list[[2]][2],my_list[[3]][2])
new_list


```


* use the typeof() function to confirm the underlying data types of each component in this list
```{r}
typeof(new_list[[1]][1]) # double
typeof(new_list[[2]][1]) # logical
typeof(new_list[[3]][1]) # character
```

* combine the underlying elements from the new list into a single atomic vector with the c() function.
```{r}
single_vec <- c(new_list[[1]], new_list[[2]], new_list[[3]])
single_vec
```


* what is the data type of this vector?

```{r}
typeof(single_vec) # character
```

## Question #3

Create a data frame with the two variables (= columns) and 26 cases (= rows) below:
```{r}
dframe <- data.frame(matrix(ncol = 2, nrow = 26))
dframe

```

* call the first variable ```my_unis``` and fill it with 26 random uniform values from 0 to 10
```{r}
names(dframe)[1] <- "my_unis"
dframe$my_unis <- runif(26, min=0, max=10)
```

* call the second variable ```my_letters``` and fill it with 26 capital letters in random order.
```{r}
names(dframe)[2] <- "my_letters"
dframe[2]

dframe$my_letters <- sample(LETTERS[1:26])
dframe
```

* for the first variable, use a single line of code in R to select 4 random rows and replace the numerical values in those rows with ```NA```.

```{r}

dframe[sample(nrow(dframe), 4), ] <- "NA"
dframe

```

* for the first variable, write a single line of R code to identify which rows have the missing values.
```{r}
which(dframe$my_unis == 'NA')
```

* re-order the entire data frame to arrange the second variable in alphabetical order
```{r}
ordered_dframe <- dframe[order(dframe$my_letters), ]
```
* calculate the column mean for the first variable.
```{r}
ordered_dframe$my_unis <- as.numeric(ordered_dframe$my_unis)
ordered_dframe
mean(ordered_dframe$my_unis, na.rm = TRUE)

```


