---
title: "Homework_06"
author: "Madelynn Edwards"
date: "2024-02-21"
output: html_document
---

# Simulating and Fitting Data Distributions
#### Fitting data distribution curves to a dataset that has intensity of earthworm invasion (biomass) and corresponding macro-arthropod abundance.

#### The simulations will be using the abundance data (Number of individuals per m^2)

```{r warning=FALSE, message=FALSE}
library(ggplot2) # for graphics
library(MASS) # for maximum likelihood estimation

# Read in data vector
z <- read.csv("plotdata_macro_constrained_20210105.csv",header=TRUE,sep=",")
str(z)
summary(z)

# Plot histogram of data

p1 <- ggplot(data=z, aes(x=abundance.m2, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)

# Add empirical density curve
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)

# Get maximum likelihood parameters for normal
normPars <- fitdistr(z$abundance.m2,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] 

# note structure of getting a named attribute
# Plot normal prob. density
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$abundance.m2),len=length(z$abundance.m2))

stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$abundance.m2), args = list(mean = meanML, sd = sdML))
p1 + stat

# Plot exponential prob. density
expoPars <- fitdistr(z$abundance.m2,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$abundance.m2), args = list(rate=rateML))
p1 + stat + stat2

# Plot uniform prob. density
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$abundance.m2), args = list(min=min(z$abundance.m2), max=max(z$abundance.m2)))
p1 + stat + stat2 + stat3

# plot gamma prob. density - this one fits best!
gammaPars <- fitdistr(z$abundance.m2,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$abundance.m2), args = list(shape=shapeML, rate=rateML))
p1 + stat + stat2 + stat3 + stat4

# Plot beta prob density
pSpecial <- ggplot(data=z, aes(x=abundance.m2/(max(abundance.m2 + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$abundance.m2/max(z$abundance.m2 + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$abundance.m2), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial

```

#### Simulate a new data set. Using the best-fitting distribution, go back to the code and get the maximum likelihood parameters. Use those to simulate a new data set, with the same length as your original vector, and plot that in a histogram and add the probability density curve. Right below that, generate a fresh histogram plot of the original data, and also include the probability density curve.

```{r warning=FALSE, message=FALSE}
zsim <- rgamma(n=80, shape=3.6329359, rate = 0.0415673)
zsim <- data.frame(1:80,zsim)
names(zsim) <- list("ID","myVar")
zsim <- zsim[zsim$myVar>0,]
str(zsim)
summary(zsim$myVar)


psim <- ggplot(data=zsim, aes(x=myVar, y=after_stat(density))) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(psim)
print(p1)
```


#### How do the two histogram profiles compare? Do you think the model is doing a good job of simulating realistic data that match your original measurements? Why or why not?

The two histogram profiles are not drastically different but there are noticeable differences between the simulated data and the original data. The simulated data using the gamma distribution parameters is a little bit more uniform in spread and has less skewness compared to the original data

#### If you have entered a large data frame with many columns, try running all of the code on a different variable to see how the simulation performs.

Insect biomass follows similar distribution as insect abundance (gamma)

Shannon diversity follows the normal distribution

Earthworm biomass seems to fit the beta distribution the best

For species richness, the normal distribution fits best
